\documentclass{article}
\usepackage{soto-homework}

\newif\ifsolution
\input{solution}

\begin{document}

\chead{ENSP 202 Homework 4}
\chead{Due Date:  Mon 24 Feb 2014}
\hrule
\vspace{10pt}

For all the following examples, explain your strategy in a few
sentences, and clearly show your work including intermediate steps.
Solutions that only provide the answer without any supporting
explanation or calculations will not receive credit.  It should be
clear from your work that you understand the material.

{\tiny Last modified: \today}

\problem{Estimations}

These estimations include areas and volumes.

\subproblem Guesstimation Problem 3.1

\solution{
The first part of the problem assumes that everyone stands together as
if they were at a concert.

$$
\ufrac{7 \cdot 10^9}{people}{}{}
\cdot
\ufrac{}{square meter}{6}{people}
\cdot
\ufrac{}{square kilometer}{10^6 }{square meter}
= 1200 square kilometers
$$

Assuming everyone is in a family of four and lives on a 30 meter by 30
meter house lot.

$$
\ufrac{7 \cdot 10^9}{ people}{}{}
\cdot
\ufrac{}{family}{4}{people}
\ufrac{900 }{square meters}{}{family}
= 1.6 \cdot 10^6 square meters
$$

}


\subproblem Guesstimation Problem 3.5

\solution{

We start by estimating the volume of the dome using a half a sphere as
an estimation.

$$ V =
\frac{1}{2}
\frac{4\pi}{3}
r^3
$$

$$ V =
\frac{1}{2}
\frac{4\pi}{3}
(16 meters)^3
= 6000 m^3
$$

$$
6000 m^3
\cdot
\ufrac{264}{gallons}{}{m$^3$}
\cdot
\ufrac{1}{minute}{3}{gallons}
= 5 \cdot 10^5 minutes
$$

}

\subproblem Guesstimation Problem 3.8

\solution{

About $ 3 \cdot 10^8$ cubic meters per year.
}

\subproblem Guesstimation Problem 3.9

\solution{

In the previous problem we estimated a yearly trash production of $3
\cdot 10^8$ cubic meters per year.

If we stack the trash 30 meters high, this will require an area of
$$\ufrac{3 \cdot 10^8}{cubic meters}{30}{meters} = 10^7 square meters$$
}

\subproblem
How many students could comfortably sunbathe on the grass between Darwin
and Stevenson?

\solution{
What is the area of the grass between the two buildings?
How much area does a student need to lay out comfortably?

According to Google maps, the grassy area is 28 meters by 40 meters.

This gives an area of $28 \cdot 40 =$ 1120 square meters.  A person is
approximately 2 meters tall and a towel is approximately 1 meter wide.
This means each person will need about 2 square meters of space.

$$ \ufrac{1120}{square meters}{}{} \cdot \ufrac{1}{person}{2}{square meters} =
560\ \textrm{persons}$$
This
}

\problem{Review}

\subproblem Binary Scientific Notation.  This number is in base 2: $101
\cdot 10^{110}$  What is this number in base 10?  (Break this up into
pieces that you know how to do.  You have all the tools you need to
figure this out.)  Be sure you understand what $10^x$ means here.

\solution{
One key part is recognizing that $10^x$ really means $2^x$.

Binary:
$$101 \cdot 10^{110} $$

$$ 101_2 = 5_{10}$$
$$ 10_2 = 2_{10}$$
$$ 110_2 = 6_{10}$$

Base 10:
$$ 5 \cdot 2^{6} $$
$$ 5 \cdot 64 $$
$$ 320 $$
}

\subproblem
Using $a \cdot c \cdot 10^{b+d}$ and showing your work, multiply
$4 \cdot 10^7$ by $2 \cdot 10^{-5}$.


\solution{
$$ 4 \cdot 2 \cdot 10^{7 + (-5)} = 8 \cdot 10^2 $$
}

\subproblem
Using $\frac{a}{c} \cdot 10^{b-d}$ and showing your work, calculate
$4 \cdot 10^7 \div 2 \cdot 10^{-5}$.

\solution{

$$ \frac{4}{2} \cdot 10^{7 - (-5)} = 2 \cdot 10^{12} $$

}

\problem{Time Spent}
Please estimate the amount of time you spent on this homework.





\end{document}
