\documentclass[12pt, oneside]{article}
\usepackage{soto-quiz}
\usepackage{tikz}
\usepackage{fancyhdr}

\newif\ifsolution
\input{solution}

\begin{document}

\lhead{ENSP 202}
\chead{Exercises 2}
\rhead{Monday 03 Mar 2014}
\lfoot{}
\cfoot{}
\rfoot{}
\pagestyle{fancy}

\namebox

\problem{}

Use this triangle for the following problem.

\begin{center}
\begin{tikzpicture}
% triangle
\draw (0,0) -- (8,0) -- (8,4) -- (0,0);
% angle label
\draw	(0.8,0.25) node[anchor=west] {$\theta$};
\end{tikzpicture}
\end{center}

Estimate the tangent of this angle and briefly explain your method.

\solution{
Using my pinky finger as an approximation for a centimeter, I measure
that the triangle is about 8cm wide and 4cm tall.  The tangent is the
height of a right triangle divided by the length, so the tangent is
about 0.5.}

\vspace{1cm}

Estimate the angle in degrees and briefly explain your method.

\solution{
I estimate that the angle is about 30 degrees.  I estimate that three of
the wedges formed by the triangle can fit in a 90 degree wedge.

This is very close to the 26.6 degrees we get from taking the inverse
tangent (or arctangent) of 0.5.

$$ \arctan 0.5 = 26.6\ degrees $$
$$ \tan 26.6\ degrees = 0.5 $$
}

\vspace{1cm}


\problem{}

Estimate the area of this ellipse in square centimeters.

\begin{center}
\begin{tikzpicture}
% ellipse
\draw (0,0) ellipse (4 and 2);
\end{tikzpicture}
\end{center}

\solution{
Using my pinky again, this ellipse is approximately 8cm by 4cm.  If it
were a rectangle, it would cover 32 cm$^2$ in area.  However, I multiply
this by about 70\% since an ellipse doesn't cover the whole rectangular
area.  From this I estimate about 21 cm$^2$.

The exact answer is $\pi \cdot 4 \cdot 2 = 25.1$
}

\end{document}

